An efficient contact value approach to approximating the pair distribution function of the inhomogeneous hard-sphere fluid

### 2014-07-14 papers/pair-correlation/paper.tex
added citations after widely used SAFT-VR Tested on knightley. Took 1:09.45.

\fixme{
  TO DO:
  \begin{itemize}
  \item \davidsays{Rewrite CVA-S section to reflect new fit.  Still working.}
  \item \pahosays{Make certain when proof-reading that other prior
    formulations cannot work in ``separable'' way.}
  \item \davidsays{Be sure that we explain the fitted data and fitting
    process well.  Still working.}
  \item \jeffsays{I will repeatedly read what you guys write!}
  \end{itemize}
}

@@

In this paper, we introduce a new contact value approach (CVA) for
the hard-sphere pair distribution function which is suitable for use
in the creation of classical density functionals based on
thermodynamic perturbation theory. We then demonstrate a more
computationally efficient separable version of the contact
value approach (CVA-S)\fixme{this needs editing}, which is based on a fit to the radial
distribution function that is separable in a way that will enable
efficient evaluation of the integral in Eq.~\ref{eq:mean-field}.

@@

\section{Contact value approach}
In the approaches for the pair distribution function
mentioned above, the radial distribution function used in the
approximation was dependent upon the density averaged over some
volume.  We seek to achieve greater accuracy by making use of a
function dependent upon our averaged $g_{\sigma}(\rr)$ discussed
above, which holds more information about an inhomogeneous system than
does a simple convolution of the density.
%
We construct the CVA with the average of two radial distribution
functions, evaluated at the distance between the two points, that are
themselves functions of the averaged pair distribution function at contact
$g_{\sigma}(\rr)$ evaluated at the two points:
%
%% We therefore seek a function form
%% for our radial distribution that is dependent upon $g_{\sigma}(\rr)$
%% and that will approach the correct solution in the homogeneous limit
%% and also in the limit of contact distance between the two points.
\begin{align}
  g^{(2)}(\rr_1,\rr_2) = \frac{g(r_{12}, g_\sigma(\rr_1)) +
    g(r_{12}, g_\sigma(\rr_2))}{2}. \label{eq:g2-our-mean}
\end{align}
This Contact Value Approach (CVA) for $g^{(2)}(\rr_1,\rr_2)$ is
constructed to reproduce the exact value for the integral:
\begin{align}
  F_1^{\text{contact}} &= \tfrac12 \iint
  g^{(2)}_{HS}(\rr_1,\rr_2)n(\rr_1)n(\rr_2)\delta(|\rr_1-\rr_2|-\sigma)
  d\rr_1d\rr_2
  \label{eq:mean-field-contact}
\end{align}
which is the mean-field correction to the free energy (see
Eq.~\ref{eq:mean-field}) for a purely contact interaction.

@@

We begin by examining the lowest-order correction in TPT, given by
Eq.~\ref{eq:mean-field}.  The two terms that are averaged in
Eq.~\ref{eq:g2-our-mean} give equal contributions to the integral:
\begin{align}
  F_1^{\textit{CVA}} &= \tfrac12 \!\! \iint \!\!
  g(r_{12},g_\sigma(\rr_2))n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|)
  d\rr_1d\rr_2
\end{align}
When we introduce the separable form for $g(r_{12},g_\sigma)$ we can
further simplify this integral as\fixme{CVA-S???}
\begin{multline}
  F_1^{\textit{CVA-S}} =
  \sum_i \tfrac12 \!\! \iint \!\! \bigg( n(\rr_1)  \bigg) \\
                             \bigg( a_i(r_{12})\Phi(r_{12}) \bigg) \\
                             \bigg( b_i(g_\sigma(\rr_2))n(\rr_2) \bigg)
  d\rr_1d\rr_2
\end{multline}
where the functional is written as a summation of integrals, each of
which is an integral of simple convolution in three dimensions.  Thus,
each of these integrals may be computed in $\mathcal{O}(N\log N)$
time, which is the same scaling behavior that is common to the most
common and accurate classical density functional theories, most
notably Fundamental Measure Theory (FMT)~\cite{rosenfeld1989free,
  roth2002whitebear}.

@@

We have performed a separable fit to the Monte-Carlo simulated radial
distribution function.  We chose to focus our fit on the range of
distances $r_{12} \le \maxrfit R$.  This range of distances is relevant to
the widely used \fixme{add citations?}  Statistical Associating Fluid
Theory of Variable Range (SAFT-VR) free energy with square-well
dispersive attraction developed by Gil-Villegas \emph{et
  al.}~\cite{gil1997statistical}.  For comparison, in
Fig.~\ref{fig:radial-distribution} we plot our fit, Monte-Carlo data,
and the radial distribution function of Gil-Villegas \emph{et al.},
which we have extracted from their approximation for the first term in
the dispersion free energy given by Eq.~\ref{eq:mean-field}.

@@

For ease of implementation and future extension to larger radii, we
chose to express our fit with a fourth-order polynomial.  We constrain
our functional form for $g_S(r, g_\sigma)$ to match the value
$g_\sigma$ at contact and to reproduce $g(r)=1$ in the low-density
($g_\sigma =1$) limit.  Incorporating these constraints we have the
functional form:
\begin{align}
  g_S(r,g_\sigma) &=
  g_\sigma + \sum_{i=1}^{4} \sum_{j=1}^{4} \kappa_{ij} (g_\sigma - 1)^i
  \left(\tfrac{r}{\sigma}-1\right)^j.
  \label{eq:fit-form}
\end{align}
where the matrix $\kappa_{ij}$ is determined from our least-squares
fit to Monte Carlo data for the radial distribution function, over the
range $2R \le r \le \maxrfit R$, and for packing fractions $\eta \le
0.45$.  The resulting parameters are displayed in
Table~\ref{tab:kappa}.  The maximum error in $g_S(r)$ within this
range is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
\rmaxerr R$.  Fig.~\ref{fig:radial-distribution} displays
our approximation at just under half of the densities that were
included in the fit.

\newcommand{\plotcomp}[1]{The top halves of these figures show the
  results of Monte Carlo simulations, while the bottom halves show the
  \fixme{CVA}CVA-S.  On the right are plots of #1 on the paths illustrated in the
  figures to the left.  These plots compare the CVA-S (blue solid line) and CVA
  (cyan dotted line) with Monte Carlo results (black circle)
  and the results of Sokolowski and Fischer
  (red dashed line)~\cite{sokolowski1992role}, and those of Fischer and Methfessel
  (green dot-dashed line)~\cite{fischer1980born}.  The latter is only plotted at
  contact, where it is defined}

@@

We begin by examining the pair distribution function near a hard wall,
with a focus on the case where one of the two spheres is in contact
with the hard wall.  Figures~\ref{fig:pair-distribution}a
and~\ref{fig:pair-distribution}c compare the results of the \fixme{CVA}CVA-S with
Monte Carlo simulations at packing fractions of 0.1 and 0.3
respectively. We see reasonable agreement at the lower density, with
some reduced oscillations at larger distances, and a flatter angular
dependence when the two spheres are in contact.  At the higher
density, we see significant structure developing in the simulation
results that is not reflected in our approximation.

\begin{figure*}
  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-contact-3.pdf}
  \vspace{-0.7cm}
\\
  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
  \vspace{-0.7cm}
  \caption{The triplet distribution function
    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
    $\rr_1$ and $\rr_2$ are in contact (a,b) and when $\rr_1$ and
    $\rr_2$ are separated by a distance $2.1\sigma$ (c,d). On the left
    are 2D plots of $g^{(3)}(\rr_1,\rr_2,\rr_3)$ as $\rr_3$
    varies. %\plotcomp{$g^{(3)}(\rr_1,\rr_2,\rr_3)$}.
%
    The top halves of these figures show the results of Monte Carlo
    simulations, while the bottom halves show the \fixme{CVA}CVA-S.  On the right
    are plots of $g^{(3)}(\rr_1,\rr_2,\rr_3)$ on the paths illustrated
    in the figures to the left.
%
    We also plot these curves along a left-right mirror image of this
    path.  The data for the right-hand paths (as shown in the 2D
    images) are marked with right-pointing triangles, while the
    left-hand paths are marked with left-pointing triangles.
%
    %% The line styles are the same as in
    %% Fig.~\ref{fig:pair-distribution}.
  }
  \label{fig:triplet-contact-distribution}
\end{figure*}

Figures~\ref{fig:pair-distribution}b and~\ref{fig:pair-distribution}d
show the pair distribution function as plotted along paths illustrated
in Figures~\ref{fig:pair-distribution}a
and~\ref{fig:pair-distribution}c.  These plots compare the CVA and
\fixme{CVA}CVA-S with Monte Carlo results, as well as the approximations of
Sokolowski and Fischer~\cite{sokolowski1992role} and of Fischer and
Methfessel~\cite{fischer1980born} at the same packing fractions of 0.1 and 0.3.
The approach of Fischer and Methfessel is only defined when the two
spheres are in contact, and is therefore only plotted on that segment
of the path.  As an input to both previous approximations and to the
CVA, we use the hard sphere radial distribution function found with
Monte Carlo simulation, interpolated as necessary. We find that both
previous approximations to the pair distribution function predict the
qualitative angular dependence of the pair distribution function at
contact better than this work.  However, in each case the pair
distribution function has a systematic error at contact---either too
high or too low.  While our approximation is smoother than either
of the existing approaches, its errors will have a tendency to cancel
when used in a perturbation expansion.  At higher densities, the
approximation of Fischer and Methfessel requires evaluating the radial
distribution function at densities significantly higher than the
freezing density, which poses numerical difficulties when using the
radial distribution function from simulation.  When the two points
$\rr_1$ and $\rr_2$ are both
more than a radius away from contact, we find that any of these
approaches gives a reasonable prediction, with the \fixme{CVA}CVA-S
underestimating the oscillations in $g^{(2)}$, as expected based on
the fit in Fig.~\ref{fig:radial-distribution}.

@@

Figures~\ref{fig:triplet-contact-distribution}a
and~\ref{fig:triplet-contact-distribution}c compare the triplet
distribution function at a packing fraction of 0.3 computed using the
\fixme{CVA}CVA-S with results from Monte Carlo simulations. In
Figure~\ref{fig:triplet-contact-distribution}a the spheres at $\rr_1$
and $\rr_2$ are in contact; in
Figure~\ref{fig:triplet-contact-distribution}c they are spaced so that
a third sphere can just fit between them; and in both figures $\rr_3$
is varied. The test-particle position for the CVA-S in each case is
$\rr_1$, which is on the left-hand side of the figure. We see
reasonable agreement, with fewer oscillations in the CVA-S at larger
distances. Also, the Monte Carlo results have the expected left-right
symmetry, while the CVA-S has an asymmetry introduced with
the test particle due to errors in the pair distribution
function. Most notably, the CVA-S has less angular dependence at
contact with the test particle than it does with the sphere at
$\rr_2$.

Figures~\ref{fig:triplet-contact-distribution}b
and~\ref{fig:triplet-contact-distribution}d show the triplet
distribution function as plotted along the paths illustrated in
Figures~\ref{fig:triplet-contact-distribution}a
and~\ref{fig:triplet-contact-distribution}c.  We also plot the
results along a left-right mirror image path, corresponding to
swapping $\rr_1$ and $\rr_2$. The two mirror-image paths are
distinguished by arrows (triangles) along the curves, with right-facing arrows
indicating the paths shown in
Figures~\ref{fig:triplet-contact-distribution}a and
\ref{fig:triplet-contact-distribution}c, and left-facing arrows
indicating the mirror image path.  As the work of
Fischer and Methfessel is only defined when $\rr_2$ and $\rr_3$ are in
contact, we only plot it along the
central portion of the path, which is in contact with $\rr_2$, and arrows
are omitted.
%
All methods perform similarly over this range with one exception.  The
separable \fixme{CVA}CVA-S method once again underestimates the oscillations on
the left-hand side of the curve, where $\rr_2$ and $\rr_3$ are
separated by a large distance.  On the path where $\rr_2$ and $\rr_3$
are closer, the CVA-S method compares favorably with the other
approximations.  This is consistent with
Fig.~\ref{fig:radial-distribution}, which shows that $g_S(r)$ fits
well the first peak of $g(r)$, but has fewer oscillations at large
distances.

@@

Figure~\ref{fig:dadz-delta} shows the results for the sticky
hard-sphere fluid.  The CVA and \fixme{NOT}CVA-S are constructed to produce this
result exactly, provided the averaged pair distribution function at
contact from Ref.~\citenum{schulte2012using} is exact.  As expected,
we see excellent agreement with the Monte Carlo simulation results,
while the approximations of Fischer and Sokolowski each show
deviations near the interface.  Figure~\ref{fig:dadz-square-well}
shows the same curve from Eq.~\ref{eq:da1} for the square-well fluid.
This potential stresses the CVA-S, since the edge of the attraction
occurs at the dip in the radial distribution function at this density,
with poor results for the bulk fluid.  We see that over these length
scales, both the CVA and Sokolowski's approximation are very accurate.
Finally, in the case of the power-law attraction all approximations
tested work very well.

>>>>

\fixme{
  TO DO:
  \begin{itemize}
  \item \davidsays{Rewrite CVA-S section to reflect new fit.  Still working.}
  \item \pahosays{Make certain when proof-reading that other prior
    formulations cannot work in ``separable'' way.}
  \item \davidsays{Be sure that we explain the fitted data and fitting
    process well.  Still working.}
  \item \jeffsays{I will repeatedly read what you guys write!}
  \item \jeffsays{still concerned about the use of CVA.  It seems
    sometimes it's used as an approach and sometimes as the resulting
    pair-dist function.}
  \end{itemize}
}

@@

\jeffsays{I've edited the following paragraph but it should be
  checked.}  In this paper, we introduce a new contact value approach
(CVA) to approximating the hard-sphere pair distribution function
which is suitable for use in the creation of classical density
functionals based on thermodynamic perturbation theory. The resulting
function is based on a fit to the radial distribution function that is
separable in a way that will enable efficient evaluation of the
integral in Eq.~\ref{eq:mean-field}.

@@

\section{Contact value approach}
In the approaches for the pair distribution function
mentioned above, the radial distribution function used in the
approximation was dependent upon the density averaged over some
volume.  We seek to achieve greater accuracy by making use of a
function dependent upon our averaged $g_{\sigma}(\rr)$ discussed
above, which holds more information about an inhomogeneous system than
does a simple convolution of the density.
%
We construct the CVA with the average of two radial distribution
functions, evaluated at the distance between the two points, that are
themselves functions of the averaged pair distribution function at contact
$g_{\sigma}(\rr)$ evaluated at the two points:
%
%% We therefore seek a function form
%% for our radial distribution that is dependent upon $g_{\sigma}(\rr)$
%% and that will approach the correct solution in the homogeneous limit
%% and also in the limit of contact distance between the two points.
\begin{align}
  g^{(2)}(\rr_1,\rr_2) = \frac{g(r_{12}, g_\sigma(\rr_1)) +
    g(r_{12}, g_\sigma(\rr_2))}{2}. \label{eq:g2-our-mean}
\end{align}
This Contact Value Approach for $g^{(2)}(\rr_1,\rr_2)$ is
constructed to reproduce the exact value for the integral:
\begin{align}
  F_1^{\text{contact}} &= \tfrac12 \iint
  g^{(2)}_{HS}(\rr_1,\rr_2)n(\rr_1)n(\rr_2)\delta(|\rr_1-\rr_2|-\sigma)
  d\rr_1d\rr_2
  \label{eq:mean-field-contact}
\end{align}
which is the mean-field correction to the free energy (see
Eq.~\ref{eq:mean-field}) for a purely contact interaction.

@@

We begin by examining the lowest-order correction in TPT, given by
Eq.~\ref{eq:mean-field}.  The two terms that are averaged in
Eq.~\ref{eq:g2-our-mean} give equal contributions to the integral:
\begin{align}
  F_1^{\textit{CVA}} &= \tfrac12 \!\! \iint \!\!
  g(r_{12},g_\sigma(\rr_2))n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|)
  d\rr_1d\rr_2
\end{align}
When we introduce the separable form for $g(r_{12},g_\sigma)$ we can
further simplify this integral as
\begin{multline}
  F_1^{\textit{CVA-S}} =
  \sum_i \tfrac12 \!\! \iint \!\! \bigg( n(\rr_1)  \bigg) \\
                             \bigg( a_i(r_{12})\Phi(r_{12}) \bigg) \\
                             \bigg( b_i(g_\sigma(\rr_2))n(\rr_2) \bigg)
  d\rr_1d\rr_2
\end{multline}
where the functional is written as a summation of integrals, each of
which is an integral of simple convolution in three dimensions.  Thus,
each of these integrals may be computed in $\mathcal{O}(N\log N)$
time, which is the same scaling behavior that is common to the most
common and accurate classical density functional theories, most
notably Fundamental Measure Theory (FMT)~\cite{rosenfeld1989free,
  roth2002whitebear}.

@@

Having settled on the basic structure of our function, we futher
refine it by performing a separable fit to the Monte-Carlo simulated
radial distribution function.  We chose to focus our fit on the range
of distances $r_{12} \le \maxrfit R$.  This range of distances is
relevant to the widely used
\cite{chapman1989saft,muller2001molecular,tan2008recent} Statistical
Associating Fluid Theory of Variable Range (SAFT-VR) free energy with
square-well dispersive attraction developed by Gil-Villegas \emph{et
  al.}~\cite{gil1997statistical}.  For comparison, in
Fig.~\ref{fig:radial-distribution} we plot our fit, Monte-Carlo data,
and the radial distribution function of Gil-Villegas \emph{et al.},
which we have extracted from their approximation for the first term in
the dispersion free energy given by Eq.~\ref{eq:mean-field}.

@@

For ease of implementation and future extension to larger radii, we
chose to express our fit with a fourth-order polynomial.  We constrain
our functional form for $g_S(r, g_\sigma)$ to match the value
$g_\sigma$ at contact and to reproduce the $g(r)=1$ result of the
low-density limit.  Incorporating these constraints we have the
functional form:
\begin{align}
  g_S(r,g_\sigma) &=
  g_\sigma + \sum_{i=1}^{4} \sum_{j=1}^{4} \kappa_{ij} (g_\sigma - 1)^i
  \left(\tfrac{r}{\sigma}-1\right)^j.
  \label{eq:fit-form}
\end{align}
where the matrix $\kappa_{ij}$ is determined from our least-squares
fit to Monte Carlo data for the radial distribution function, over the
range $2R \le r \le \maxrfit R$, and for packing fractions $\eta \le
0.45$.  The resulting parameters are displayed in
Table~\ref{tab:kappa}.  The maximum error in $g_S(r)$ within this
range is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
\rmaxerr R$.  Fig.~\ref{fig:radial-distribution} displays
our approximation at just under half of the densities that were
included in the fit.

\newcommand{\plotcomp}[1]{\jeffsays{there's no green dash label in the
    legend above.  Do we want the green dashes?}The top halves of
  these figures show the results of Monte Carlo simulations, while the
  bottom halves show the CVA.  On the right are plots of #1 on the
  paths illustrated in the figures to the left.  These plots compare
  the CVA (blue solid line), Monte Carlo results (black circles), the
  results of Sokolowski and Fischer (red dashed
  line)~\cite{sokolowski1992role}, and those of Fischer and Methfessel
  (green dot-dashed line)~\cite{fischer1980born}.  The latter is only
  plotted at contact, where it is defined}

@@

We begin by examining the pair distribution function near a hard wall,
with a focus on the case where one of the two spheres is in contact
with the hard wall.  Figures~\ref{fig:pair-distribution}a
and~\ref{fig:pair-distribution}c compare the results of the CVA with
Monte Carlo simulations at packing fractions of 0.1 and 0.3
respectively. We see reasonable agreement at the lower density, with
some reduced oscillations at larger distances, and a flatter angular
dependence when the two spheres are in contact.  At the higher
density, we see significant structure developing in the simulation
results that is not reflected in our approximation.

\begin{figure*}
  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-contact-3.pdf}
  \vspace{-0.7cm}
\\
  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
  \vspace{-0.7cm}
  \caption{The triplet distribution function
    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
    $\rr_1$ and $\rr_2$ are in contact (a,b) and when $\rr_1$ and
    $\rr_2$ are separated by a distance $2.1\sigma$ (c,d). On the left
    are 2D plots of $g^{(3)}(\rr_1,\rr_2,\rr_3)$ as $\rr_3$
    varies. %\plotcomp{$g^{(3)}(\rr_1,\rr_2,\rr_3)$}.
%
    The top halves of these figures show the results of Monte Carlo
    simulations, while the bottom halves show the CVA.  On the right
    are plots of $g^{(3)}(\rr_1,\rr_2,\rr_3)$ on the paths illustrated
    in the figures to the left.
%
    We also plot these curves along a left-right mirror image of this
    path.  The data for the right-hand paths (as shown in the 2D
    images) are marked with right-pointing triangles, while the
    left-hand paths are marked with left-pointing triangles.
%
    %% The line styles are the same as in
    %% Fig.~\ref{fig:pair-distribution}.
  }
  \label{fig:triplet-contact-distribution}
\end{figure*}

Figures~\ref{fig:pair-distribution}b and~\ref{fig:pair-distribution}d
show the pair distribution function as plotted along paths illustrated
in Figures~\ref{fig:pair-distribution}a
and~\ref{fig:pair-distribution}c.  These plots compare the CVA with
Monte Carlo results, as well as the approximations of Sokolowski and
Fischer~\cite{sokolowski1992role} and of Fischer and
Methfessel~\cite{fischer1980born} at the same packing fractions of 0.1
and 0.3.  The approach of Fischer and Methfessel is only defined when
the two spheres are in contact, and is therefore only plotted on that
segment of the path.  As an input to both previous approximations and
to the CVA, we use the hard sphere radial distribution function found
with Monte Carlo simulation, interpolated as necessary. We find that
both previous approximations to the pair distribution function predict
the qualitative angular dependence of the pair distribution function
at contact better than this work.  However, in each case the pair
distribution function has a systematic error at contact---either too
high or too low.  While our approximation is smoother than either of
the existing approaches, its errors will have a tendency to cancel
when used in a perturbation expansion.  At higher densities, the
approximation of Fischer and Methfessel requires evaluating the radial
distribution function at densities significantly higher than the
freezing density, which poses numerical difficulties when using the
radial distribution function from simulation.  When the two points
$\rr_1$ and $\rr_2$ are both more than a radius away from contact, we
find that any of these approaches gives a reasonable prediction, with
the CVA underestimating the oscillations in $g^{(2)}$, as
expected based on the fit in Fig.~\ref{fig:radial-distribution}.

@@

Figures~\ref{fig:triplet-contact-distribution}a
and~\ref{fig:triplet-contact-distribution}c compare the triplet
distribution function at a packing fraction of 0.3 computed using the
CVA with results from Monte Carlo simulations. In
Figure~\ref{fig:triplet-contact-distribution}a the spheres at $\rr_1$
and $\rr_2$ are in contact; in
Figure~\ref{fig:triplet-contact-distribution}c they are spaced so that
a third sphere can just fit between them; and in both figures $\rr_3$
is varied. The test-particle position for the CVA in each case is
$\rr_1$, which is on the left-hand side of the figure. We see
reasonable agreement, with fewer oscillations in the CVA at larger
distances. Also, the Monte Carlo results have the expected left-right
symmetry, while the CVA has an asymmetry introduced with the test
particle due to errors in the pair distribution function. Most
notably, the CVA has less angular dependence at contact with the
test particle than it does with the sphere at $\rr_2$.

Figures~\ref{fig:triplet-contact-distribution}b
and~\ref{fig:triplet-contact-distribution}d show the triplet
distribution function as plotted along the paths illustrated in
Figures~\ref{fig:triplet-contact-distribution}a
and~\ref{fig:triplet-contact-distribution}c.  We also plot the
results along a left-right mirror image path, corresponding to
swapping $\rr_1$ and $\rr_2$. The two mirror-image paths are
distinguished by arrows (triangles) along the curves, with right-facing arrows
indicating the paths shown in
Figures~\ref{fig:triplet-contact-distribution}a and
\ref{fig:triplet-contact-distribution}c, and left-facing arrows
indicating the mirror image path.  As the work of
Fischer and Methfessel is only defined when $\rr_2$ and $\rr_3$ are in
contact, we only plot it along the
central portion of the path, which is in contact with $\rr_2$, and arrows
are omitted.
%
All methods perform similarly over this range with one exception.  The
CVA method once again underestimates the oscillations on the left-hand
side of the curve, where $\rr_2$ and $\rr_3$ are separated by a large
distance.  On the path where $\rr_2$ and $\rr_3$ are closer, the CVA
method compares favorably with the other approximations.  This is
consistent with Fig.~\ref{fig:radial-distribution}, which shows that
$g_S(r)$ fits well the first peak of $g(r)$, but has fewer
oscillations at large distances.

@@

Figure~\ref{fig:dadz-delta} shows the results for the sticky
hard-sphere fluid.  The CVA is constructed to produce this
result exactly, provided the averaged pair distribution function at
contact from Ref.~\citenum{schulte2012using} is exact.  As expected,
we see excellent agreement with the Monte Carlo simulation results,
while the approximations of Fischer and Sokolowski each show
deviations near the interface.  Figure~\ref{fig:dadz-square-well}
shows the same curve from Eq.~\ref{eq:da1} for the square-well fluid.
This potential stresses the CVA, since the edge of the attraction
occurs at the dip in the radial distribution function at this density,
with poor results for the bulk fluid.  We see that over these length
scales, both the CVA and Sokolowski's approximation are very accurate.
Finally, in the case of the power-law attraction all approximations
tested work very well.
